Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring

### 2014-08-20 micro14/rebuttal.txt

Review-A

-How is metadata dependence graph encoded and processed?
The full graph structure does not need to be stored. The dataflow engine
dynamically propagates the information about whether metadata is null or
invalid. Source-only dropping is enabled by only making dropping decisions on
explicit instructions which modify metadata. This does not require information
about the full dependence graph to identify.

Review-B

-Previous work has removed null-metadata (e.g., FlexiTaint).
As we mention in the paper, we are not the first work to filter out null
metadata. However, we show how a single dataflow engine can be used to provide
both this filtering mechanism as well a dropping mechanism to further reduce
overheads.

-FlexCore allows an instruction type to be not forwarded or not forwarded if
the FIFO is full.
Our implementation also filters out instruction types which are not
relevant to the monitor. Not forwarding instructions if the FIFO is full can
lead to false positives which FlexCore does not prevent.

-Baseline overheads are high. In addition, FlexCore shows much lower overheads.
One reason for our high baseline overheads is due to the need to dynamically
allocate memory space for metadata due to the large metadata we are using.
FlexCore shows much lower overheads by performing monitoring on a
reconfigurable fabric. This requires dedicated on-chip reconfigurable
fabric for this purpose. Our evaluation focuses on a core-based monitor but our
techniques could also be applied to FlexCore to further reduce its overheads.

-Information flow tracking can have false positives and negatives (e.g.,
implicit flows). Is the effect of these examined?
In this work we have focused on the false positives and negatives that arise
from dropping a portion of the information flow tracking operations. We have
not looked at the inherent false positives and negatives that arise with
specific information flow tracking schemes, such as handling implicit flows.

-Why use extra HW + extra core as opposed to a programmable HW unit?
We focused our evaluation on a core-based monitor as this has greater
flexibility and requires less hardware changes to implement.
However, our techniques could also be applied to hardware monitors such as
FlexiTaint and FlexCore to enable reduced overheads by filtering and
dropping monitoring events.

Review-C

-Why choose security monitoring instead of applications like profiling which
 can tolerate partial coverage?
Profiling would also serve as a good application for our system and we
could study such a scheme on our system in the future. We chose to focus
on security monitoring as the issues with targeting security monitoring
were a superset of those for most profiling monitors (e.g., dealing with false
positives).

Review-D

-Core idea of Section 2 (monitoring) is same as Heapmon.
Section 2 describes our baseline run-time monitoring architecture.  Our work
extends previous work on run-time monitoring, such as Heapmon and others, by
filtering out monitoring on null metadata and dropping monitoring events to
further reduce overheads.

-Claiming false negatives can be ignored undermines the soundness of the
 monitoring. Can you afford for tainted variable to be used by untrusted
 function?
Although full monitoring is always desired, if the performance overheads
of full monitoring cannot be tolerated, then false negatives are a
necessary trade off. The alternative would be to completely disable
monitoring (i.e., 100% false negatives).

-Claim that work can emulate all prior works on memory debugging and control
 flow integrity check. How can your scheme emulate every single one of them?
As the monitor is implemented in software on a core, it can support a wide
range of monitoring operations. Specifically, any monitoring technique
which performs operations in response to instruction executions should be
implementable.

Review-E

-How much additional benefit does data-flow filtering provide over
 instruction-type filtering (FlexCore)?
Like FlexCore, our system also filters out monitoring for instruction
types which are not relevant to a monitoring technique. The results we
show are the additional benefit of our technique on top of
instruction-type filtering.

-Baseline selected based on what runs on simulator instead of
 state-of-the-art (dynamic allocation due to memory issues with simulator). 
Monitoring tags can require a large memory space. This is a problem not
unique to our baseline, though it is not always addressed. We have chosen to
deal with this by performing dynamic allocation of memory for metadata rather
than statically allocating a large memory space for metadata (up to 4 GB).
This comes at a trade-off in overheads. UMC, which has smaller metadata, was
implemented by statically allocating the metadata memory and thus does not
see as high overheads as BC and DIFT. A dedicated metadata TLB could be used
to reduce these overheads.

-Processor with hardware support for information flow tracking and support
 for configurable policies can be re-purposed to provide same functionality as
 the technique proposed in the paper. 
Our technique for using a dataflow engine for filtering and dropping metadata
can be applied to both core-based monitors and dedicated hardware monitors such
as a hardware-supported information flow tracking mechanism. For a sufficiently
configurable information flow track module, it may be possible to incorporate
this into the information flow tracking policy.  However, as far as we know, we
are the first to propose using this flow tracking mechanism to provide both
filtering and dropping of monitoring events to reduce monitoring overheads. In
addition, there are monitoring schemes that may not be possible to be mapped to
information flow tracking (e.g., control flow integrity checking, hardware
error detection, etc.) in which case a separate dataflow engine, such as the
one we describe, would be needed to enable filtering and dropping.

### 2014-08-21 micro14/rebuttal.txt

General

-How does this work compared to previous work (FlexiTaint, FlexCore, Heapmon,
etc.)?
Run-time monitoring is a well studied problem. However, our work is the first
that proposes a general hardware platform that allows for an adjustable
overhead target by trading off coverage for overhead.

@@

Review-A

-How is metadata dependence graph encoded and processed?
Sources are marked statically in the program by the compiler or programmer
annotation. The dataflow engine stores 2 bits per metadata indicating whether
it is null or invalid. It dynamically propagates this information as
instructions execute.

@@

Review-B

-How does this compare to FlexCore's dropping of events?
We ensure that false positives do not occur when events are dropped. FlexCore
does not prevent these false positives.

-Baseline overheads are high.
Our overheads are comparable to previous work such as [Newsome NDSS'05]
(24-37x) and [Chen ASID'06] (4-10x). More recent works have aggressively
optimized these monitors to reduce overheads, though these overheads could
still be considered too high to enable monitoring in certain contexts. These
optimizations could be applied to our system as well, resulting in higher
coverage at the same target overhead or comparable coverage at even lower
overheads.

-Information flow tracking can have false positives and negatives (e.g.,
implicit flows). Is the effect of these examined?
The information flow tracking scheme that we examined only tracked explicit
flows. Monitoring schemes which track implicit flows may require additional
support to ensure no false positives on our hardware platform.

-Why use extra HW + extra core as opposed to a programmable HW unit?
We evaluated our idea on a core-based monitor as this has greater flexibility
and requires less hardware changes to implement. However, our techniques could
also be applied to hardware monitors, such as FlexiTaint and FlexCore, to enable
adjustable overheads.

@@

Review-C

-Why choose security monitoring instead of applications like profiling which
 can tolerate partial coverage?
We focused on security monitoring due to the challenges that arise with false
positives for these monitoring techniques. Our platform could also be used with
profiling applications in order to enable reduced and adjustable overheads.

@@

Review-D

-Claiming false negatives can be ignored.
Although full monitoring is always desired, if the performance overheads
of full monitoring cannot be tolerated, then false negatives are a
necessary trade off. The alternative would be to completely disable
monitoring (i.e., 100% false negatives).

-How can your scheme emulate other monitoring techniques?
As the monitor is implemented in software on a core, it can support a wide
range of monitoring operations. Specifically, any instruction-grained
monitoring technique should be implementable.

@@

Review-E

-How much additional benefit does data-flow filtering provide over
 instruction-type filtering (FlexCore)?
Like FlexCore, our system also filters out monitoring for instruction
types which are not relevant to a monitoring technique. The results we
show are the additional benefit of our technique on top of
instruction-type filtering.

-Baseline selected based on what runs on simulator instead of
 state-of-the-art (dynamic allocation of memory).
Dynamic allocation of metadata memory is required for the large metadata used
by the monitors we evaluated (up to 4 GB total memory required). Static
allocation may be possible for systems with large memory space, but was not
reasonable for our target system.

-Processor with hardware support for information flow tracking can provide the
same functionality.
The main focus of our platform is to enable reduced and adjustable overheads
which has not been studied by previous works on hardware support for
information flow tracking. The monitoring techniques we evaluated could be
implemented with a sufficiently configurable information flow tracking
mechanism. However, there are other monitors (e.g., control flow integrity
checking, hardware error detection, etc.) which our technique could also be
applied to.

### 2014-08-22 micro14/rebuttal.txt

General

-How does this work compared to previous work (FlexiTaint, FlexCore, Heapmon,
etc.)?
Run-time monitoring is a well studied problem. However, our work is the first
that proposes a general hardware platform that allows for an adjustable
overhead target by trading off coverage for overhead.
While we use a core-based monitor to demonstrate techniques, our platform
is applicable to other types of monitors such as FlexCore.

@@

Review-A

-How is metadata dependence graph encoded and processed?
Sources are marked statically in the program by the compiler or programmer
annotation. The dataflow engine stores 2 bits per metadata indicating whether
it is null or invalid. It dynamically propagates this information as
instructions execute.

Review-B

-How does this compare to FlexCore's dropping of events?
We ensure that false positives do not occur when events are dropped. FlexCore
does not prevent these false positives. Our dropping is also more intelligent
selectively dropping based on the overhead target and importance.

-Baseline overheads are high.
Our overheads are comparable to previous work such as [Newsome NDSS'05]
(24-37x) and [Chen ASID'06] (4-10x). More recent works have aggressively
optimized these monitors to reduce overheads. Yet, the overhead especially
for monitors with large tags can still be significant.
The dataflow engine can be applied to the optimized monitors to further
reduce the overhead or enable coverage at a lower overhead target.

-The dataflow engine does not track implicit flows. Can this lead to
false positives or negatives?
The monitors that we examined only tracked explicit flows. As a results,
the dataflow engine only needed to track explicit flows.
Monitoring schemes which track implicit flows may require additional
support to ensure no false positives on our hardware platform.

-Why use extra HW + extra core as opposed to a programmable HW unit?
We evaluated our idea on a core-based monitor as this represents the most
flexible design and leverages multi-cores with less hardware changes.
However, our techniques could
also be applied to hardware monitors, such as FlexiTaint and FlexCore, to enable
adjustable overheads.

@@

Review-C

-Why choose security monitoring instead of applications like profiling which
 can tolerate partial coverage?
We agree that profiling is a promising application for our platform with
adjustable overhead.
We focused on monitors for program bugs because they address important problems
and represent a general case where false positives need to be handled.

@@

Review-D

-Claiming false negatives can be ignored.
Although full monitoring is always desired, if the performance overheads
of full monitoring cannot be tolerated, then false negatives are a
necessary trade off. The alternative would be to completely disable
monitoring (i.e., 100% false negatives).

-How can your scheme emulate other monitoring techniques?
As the monitor is implemented in software on a core, it can support a wide
range of monitoring operations. Specifically, any technique that monitors
a trace of instructions should be implementable.

@@

Review-E

-How much additional benefit does data-flow filtering provide over
 instruction-type filtering (FlexCore)?
Like FlexCore, our baseline system filters out monitoring for instruction
types which are not relevant to a monitoring technique. The results we
show are the additional benefit of our technique on top of
instruction-type filtering.

-Baseline selected based on what runs on simulator instead of
 state-of-the-art (dynamic allocation of memory).
We implemented dynamic metadata allocation because it is required for monitors
with large metadata (for example, 64-bits per word in bound checking or
32-bit per byte for multi-bit DIFT). We also tested static allocation for 
smaller metadata.

-Processor with hardware support for information flow tracking can provide the
same functionality.
The main focus of our platform is to enable reduced and adjustable overheads
which has not been studied by previous works on hardware support for
information flow tracking. The monitoring techniques we evaluated could be
implemented with a sufficiently configurable information flow tracking
mechanism. However, there are other monitors (e.g., control flow integrity
checking, hardware error detection, race detection, etc.) that cannot be 
implemented as DIFT but can use our technique.

### 2014-08-22 micro14/rebuttal.txt

Review-A

-How is metadata dependence graph encoded and processed?
Sources are marked statically in the program by the compiler or programmer
annotation. The dataflow engine stores 2 bits per metadata indicating whether
it is null or invalid. It dynamically propagates this information as
instructions execute.

Review-B

-How does this compare to FlexCore's dropping of events?
We ensure that false positives do not occur when events are dropped. FlexCore
does not prevent these false positives. Our dropping is also more intelligent
by selectively dropping based on the overhead target and importance.

-Baseline overheads are high.
Our overheads are comparable to previous work such as [Newsome NDSS'05]
(24-37x) and [Chen ASID'06] (4-10x). More recent works have aggressively
optimized these monitors to reduce overheads. Yet, the overhead, especially for
monitors with large tags, can still be significant. Our dataflow engine can be
applied to the optimized monitors to further reduce the overhead or enable
coverage at a lower overhead target.

@@

-How can your scheme emulate other monitoring techniques?
As the monitor is implemented in software on a core, it can support a wide
range of monitoring operations. Specifically, any technique that monitors
a trace of instructions should be implementable.

Review-E

-How much additional benefit does data-flow filtering provide over
 instruction-type filtering (FlexCore)?
Like FlexCore, our baseline system filters out monitoring for instruction
types which are not relevant to a monitoring technique. The results we
show are the additional benefit of our technique on top of
instruction-type filtering.

-Baseline selected based on what runs on simulator instead of
 state-of-the-art (dynamic allocation of memory).
We implemented dynamic metadata allocation because it is required for monitors
with large metadata (for example, 64-bits per word in bound checking or
32-bits per byte for multi-bit DIFT). We also tested static allocation for 
smaller metadata.

### 2014-09-05 evaluation.tex
Reorganized paper for HPCA 2015 submission.

\subsection{Baseline Monitoring Overheads}
% Full monitoring overheads
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_full_mon.pdf}
    \vspace{-0.2in}
    \caption{Full monitoring overheads for UMC, BC, and DIFT.}
    \label{fig:evaluation.full_mon}
    \vspace{-0.1in}
  \end{center}
\end{figure} Figure~\ref{fig:evaluation.full_mon} shows the execution times of
performing full monitoring normalized to the execution times of the benchmarks
without monitoring. In these results, no filtering or partial monitoring is
done. UMC shows normalized execution times from 4x to 11x with an average
of 7x. BC shows normalized execution times of 9-28x with an average of
18x while DIFT shows normalized execution times of 6x-17x with an average of
11x. One of the reasons for these high overheads is that our implementations of
these monitors dynamically allocate memory for new metadata. This can be an
expensive, many-cycle operation. Statically allocating memory for metadata
will reduce the execution time but requires a large upfront memory footprint
which caused issues with our simulator.

>>>>

\subsection{Baseline Monitoring Overheads}

% Full monitoring overheads
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_full_mon.pdf}
    \vspace{-0.2in}
    \caption{Full monitoring overheads for UMC, BC, and DIFT.}
    \label{fig:evaluation.full_mon}
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:evaluation.full_mon} shows the execution times of
performing full monitoring normalized to the execution times of the benchmarks
without monitoring. In these results, no filtering or partial monitoring is
done. UMC shows normalized execution times from 4x to 11x with an average
of 7x. BC shows normalized execution times of 9-28x with an average of
18x while DIFT shows normalized execution times of 6x-17x with an average of
11x. One of the reasons for these high overheads is that our implementations of
these monitors dynamically allocate memory for new metadata. This can be an
expensive, many-cycle operation. Statically allocating memory for metadata
will reduce the execution time but requires a large upfront memory footprint
which caused issues with our simulator.

### 2014-09-05 introduction.tex
Reorganized paper for HPCA 2015 submission.

% Programmable hardware monitors
These run-time monitors introduce performance, power, and energy overheads.
% Implementing run-time monitoring in software using binary instrumentation or
% other similar methods has been shown to introduce high overheads. For example,
% dynamic information flow tracking (DIFT) implemented in software suffers a 3.6x
% slowdown \cite{lift-micro06}. 
Implementing monitors in dedicated hardware has been shown to have small
performance impacts. For example, a dedicated hardware implementation of
dynamic information flow tracking (DIFT) shows performance
overheads of only 1.1\% \cite{dift-asplos04}. However, fixed hardware loses
the programmability of software, and cannot be updated or changed. It 
requires for monitors to be chosen and implemented at hardware design time.
Recent studies \cite{lba-isca08, ruwase-spaa08, paralog-asplos10, oases-sigops09} have proposed using parallel cores in a
multi-core system to perform monitoring. This allows monitoring to be done
using parallel hardware while having the flexibility and programmability of
software. However, due to the multiple cycles it may take for software to
monitor an instruction, the overheads can easily increase execution time
several fold.
% Two optimizations
In this paper, we present a hardware architecture that uses dynamic dataflow
tracking in order to reduce the overheads of programmable parallel monitoring
by filtering out unnecessary monitoring operations and by enabling partial
monitoring.

% Filtering
Recent studies have shown that filtering out monitoring checks on clean or null
metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}.
For example, for
an array bounds check, checking non-pointer accesses is unnecessary. 
In this paper, we use a hardware-based dataflow tracking engine to track and
filter out these null metadata monitoring operations.
By using the dataflow engine to track a separate set of flags
indicating null metadata, we are able to handle the common case of null
metadata without the need to read the full metadata values.

There is a limit to how effective filtering is at reducing monitoring overheads
as it only eliminates unnecessary monitoring operations. The ability
to also skip relevant monitoring operations can be used to further reduce overheads.
Although this comes at the cost of reduced monitoring coverage (i.e.,
effectiveness), this partial monitoring can still be very useful.
For example, partial
monitoring can enable a level of protection even for systems where the full
monitoring overheads are too high. This is especially true for energy- or
power-constrained systems or soft real-time systems where the monitoring
overhead should not exceed energy/power limits or real-time deadlines.
Additionally, partial monitoring can be used for debugging purposes
especially in the context of sampling-based or cooperative debugging techniques
which expect low coverage per run but use a large number of runs and users to
collect debugging information~\cite{liblit-pldi05, chilimbi-asplos04,
greathouse-cgo11}.

% % Adjustable overhead
In this paper, we show how our dataflow-guided architecture is also able to
enable partial monitoring. This is achieved by dynamically dropping monitoring
operations when overheads exceed a specified overhead budget. We prevent false
positives from partial monitoring by using the dataflow engine to track these
dropped monitoring flows. With this architecture, we investigate different
policies for deciding which events to drop for partial monitoring. These
different policies show a trade-off between closely matching the overhead
budget and the monitoring coverage achieved.

The main contributions of this paper are summarized as follows.
\begin{itemize}
  \item We propose two optimizations for reducing the overheads of monitoring: filtering null metadata and performing partial monitoring.
  \item We present a hardware architecture using a dataflow tracking engine to
  efficiently enable filtering and partial monitoring. This hardware
  architecture is designed to be generally applicable to a wide-range of run-time monitoring
  techniques.
  \item We investigate interesting trade-off decisions for partial monitoring on the proposed architecture.
\end{itemize}

@@

% Section summary
This paper is organized as follows. Section~\ref{sec:monitoring} presents the
parallel run-time monitoring model that this paper assumes.
Section~\ref{sec:optimizations} presents the high-level approach of how we aim
to reduce overheads through and Section~\ref{sec:arch} describes our
hardware architecture that supports this. Section~\ref{sec:policies}
investigates the design space for dropping decisions for partial monitoring.
Section~\ref{sec:evaluation} presents our evaluation methodology and
results. Finally, we discuss related work in Section~\ref{sec:related} and
conclude in Section~\ref{sec:conclusion}.

>>>>

% Programmable hardware monitors
These run-time monitors introduce performance, power, and energy overheads.
% Implementing run-time monitoring in software using binary instrumentation or
% other similar methods has been shown to introduce high overheads. For example,
% dynamic information flow tracking (DIFT) implemented in software suffers a 3.6x
% slowdown \cite{lift-micro06}. 
Implementing monitors in dedicated hardware has been shown to have small
performance impacts. For example, a dedicated hardware implementation of
dynamic information flow tracking (DIFT) shows performance
overheads of only 1.1\% \cite{dift-asplos04}. However, fixed hardware loses
the programmability of software, and cannot be updated or changed. It 
requires for monitors to be chosen and implemented at hardware design time.
Recent studies \cite{lba-isca08, ruwase-spaa08, paralog-asplos10, oases-sigops09} have proposed using parallel cores in a
multi-core system to perform monitoring. This allows monitoring to be done
using parallel hardware while having the flexibility and programmability of
software. However, due to the multiple cycles it may take for software to
monitor an instruction, the overheads can easily increase execution time
several fold.
% Two optimizations
Since these overheads can be prohibitive for run-time use, we propose to enable
partial monitoring in order to reduce the overheads.

Although partial monitoring comes at the cost of reduced monitoring coverage (i.e.,
effectiveness), this can still be very useful.
For example, partial
monitoring can enable a level of protection even for systems where the full
monitoring overheads are too high. This is especially true for energy- or
power-constrained systems or soft real-time systems where the monitoring
overhead should not exceed energy/power limits or real-time deadlines.
Additionally, partial monitoring can be used for debugging purposes
especially in the context of sampling-based or cooperative debugging techniques
which expect low coverage per run but use a large number of runs and users to
collect debugging information~\cite{liblit-pldi05, chilimbi-asplos04,
greathouse-cgo11}.

% Adjustable overhead
Partial monitoring with adjustable overheads is achieved by dynamically
dropping monitoring operations when overheads exceed a specified overhead
budget.  In order to prevent dropped monitoring operations from leading to a
false positive, we need to track these dropped flows.  We present a
hardware-based dataflow tracking engine that keeps track of these dropped
flows. With this architecture, we investigate different policies for deciding
which events to drop for partial monitoring. These different policies show a
trade-off between closely matching the overhead budget and the monitoring
coverage achieved.

% Filtering
Recent studies have shown that filtering out monitoring checks on clean or null
metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}.
For example, for an array bounds check, checking non-pointer accesses is
unnecessary. In addition to enabling partial monitoring, we show how a simple
extension to our dataflow engine enables this null metadata filtering.

The main contributions of this paper are summarized as follows:
\begin{itemize}
  \item We present a hardware architecture using a dataflow tracking engine to
  efficiently enable partial monitoring while preventing false positives. This
  hardware architecture is designed to be generally applicable to a wide-range
  of run-time monitoring techniques.
  \item We investigate interesting trade-offs in deciding what to skip for
  partial monitoring on the proposed architecture.
  \item We extend the dataflow engine to enable filtering null metadata to
  reduce overheads.
\end{itemize}

@@

% Section summary
This paper is organized as follows. Section~\ref{sec:monitoring} presents the
baseline parallel run-time monitoring model that this paper assumes.
Section~\ref{sec:dropping} discusses enabling partial monitoring using our dataflow tracking engine.
Section~\ref{sec:policies}
investigates the design space for dropping decisions for partial monitoring.
Section~\ref{sec:filtering} shows how the dataflow tracking engine can be extended to filter null metadata.
Section~\ref{sec:evaluation} presents our evaluation methodology and
results. Finally, we discuss related work in Section~\ref{sec:related} and
conclude in Section~\ref{sec:conclusion}.

### 2014-09-05 monitoring.tex
More edits based on Ed's comments.
Reorganized paper for HPCA 2015 submission.

@@

There are many possible monitoring schemes that can be implemented on this type
of fine-grained parallel monitoring architecture such as memory protection
\cite{mondrian-asplos02}, information flow tracking \cite{dift-asplos04,
testudo-micro08}, soft error detection \cite{argus-micro07}, data-race
detection \cite{cord-hpca06}, etc.  For example, an array bounds check (BC)
\cite{hardbound-asplos08} can be implemented in order to detect
when software attempts to read or write to a memory location outside of an
array's bounds. This can be done by associating metadata with array pointers that 
indicates the array's base (start) and bound (end) addresses. On loads or stores with the
array
pointer, the monitoring core checks that the memory address accessed is within the base and
bound addresses. In addition, this base and bound metadata
is propagated on certain ALU instructions to track the creation of new array pointers.
% One example is an uninitialized memory check (UMC) where monitoring is used to
% detect when software attempts to read from a memory location that was not
% previously initialized. This can be done by forwarding each load and store
% instruction from the main core to the monitor. For every memory location, the
% monitor keeps one bit of metadata. On a store to a memory location, the monitor
% marks the corresponding metadata bit to indicate that the memory location has
% been initialized. On a load, the monitor checks that the corresponding metadata
% bit has been previously marked as initialized.

% There are multiple options for implementing programmable parallel monitors. For
% example, the Log-Based Architecture \cite{lba-isca08} uses processor cores in a
% multi-core system as monitors. The FlexCore architecture
% \cite{flexcore-micro10} instead uses an FPGA-fabric to implement the
% monitor. The approach we describe in this paper applies to any of these
% parallel monitors. However, for experiments, we model an FPGA-based monitor
% similar to FlexCore.  The on-chip FPGA fabric is used to implement the
% ``Monitor'' block in Figure~\ref{fig:arch.overview} while the FIFO from the
% main core and metadata cache are implemented as ASICs.

>>>>

\subsection{Overheads of Run-Time Monitoring}

% Previous work overheads
\begin{table*}[tb]
  \begin{center}
    \vspace{-0.0in}
    \begin{footnotesize}
    \input{tabs/previous_overheads}
    \end{footnotesize}
    \caption{Overheads of various proposed run-time monitoring systems.}
    \vspace{-0.2in}
    \label{tab:monitoring.previous_overheads}
  \end{center}
\end{table*}

There have been numerous previous works exploring various implementations of
run-time monitoring systems. Table~\ref{tab:monitoring.previous_overheads}
summarizes some of these implementations and the average and worst-case
overheads reported. These systems vary in their hardware complexity,
flexibility, monitoring schemes, and overheads. Dedicated hardware can provide low 
overheads for specific monitoring schemes. However, more pronounced overheads are seen for 
flexible monitoring systems that implement monitoring on a processor core
\cite{lba-asid06, lba-isca08, nagarajan-interact08} or using an FPGA fabric
\cite{flexcore-micro10}. These systems allow a wide range of monitoring schemes to
be implemented on the same hardware architecture and also allow monitors to be modified and updated after deployment. However, in order to support this flexibility, they can easily incur tens of percent of
overhead in the average-case and severalfold slowdowns in the worst-case. Partial monitoring
can be a useful technique for reducing overheads for any of these systems.
In this paper, we will focus our discussion on using a parallel
processing core to implement the monitor. Our evaluation includes results
for both the core-based monitor as well as a higher performance FPGA-based monitor.

\subsection{Baseline Architecture}

@@

There are many possible monitoring schemes that can be implemented on this type
of fine-grained parallel monitoring architecture such as memory protection
\cite{mondrian-asplos02}, information flow tracking \cite{dift-asplos04,
testudo-micro08}, soft error detection \cite{argus-micro07}, data-race
detection \cite{cord-hpca06}, etc.  For example, an array bounds check (BC)
\cite{hardbound-asplos08} can be implemented in order to detect
when software attempts to read or write to a memory location outside of an
array's bounds. This can be done by associating metadata with array pointers that 
indicates the array's base (start) and bound (end) addresses. On loads or stores with the
array
pointer, the monitoring core checks that the memory address accessed is within the base and
bound addresses. In addition, this base and bound metadata
is propagated on ALU and memory instructions to track the corresponding array pointers.
% One example is an uninitialized memory check (UMC) where monitoring is used to
% detect when software attempts to read from a memory location that was not
% previously initialized. This can be done by forwarding each load and store
% instruction from the main core to the monitor. For every memory location, the
% monitor keeps one bit of metadata. On a store to a memory location, the monitor
% marks the corresponding metadata bit to indicate that the memory location has
% been initialized. On a load, the monitor checks that the corresponding metadata
% bit has been previously marked as initialized.

% There are multiple options for implementing programmable parallel monitors. For
% example, the Log-Based Architecture \cite{lba-isca08} uses processor cores in a
% multi-core system as monitors. The FlexCore architecture
% \cite{flexcore-micro10} instead uses an FPGA-fabric to implement the
% monitor. 
% 
% The approach we describe in this paper applies to any of these
% parallel monitors. However, for experiments, we model an FPGA-based monitor
% similar to FlexCore.  The on-chip FPGA fabric is used to implement the
% ``Monitor'' block in Figure~\ref{fig:arch.overview} while the FIFO from the
% main core and metadata cache are implemented as ASICs.

% Example monitor
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/example_full.pdf}
    \vspace{-0.2in}
    \caption{Example of array bounds check monitor.}
    \label{fig:monitoring.example_full}
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:monitoring.example_full} shows an example pseudo-code segment and
the corresponding monitoring operations for an array bounds check monitor. 
First, an array {\tt x} is allocated using {\tt malloc}. When this happens, the
monitor associates the array with metadata corresponding to its base and bounds
address. Next, {\tt y} is set to point to the middle of {\tt y}. The metadata
for {\tt y} is the same as the metadata {\tt x} to ensure that accesses using
{\tt y} do not exceed the bounds of {\tt x}. Finally, both pointers {\tt x} and
{\tt y} are used to perform a write. In both cases, the monitor checks whether
the address that is written to is within the bounds stored in the metadata. In
the first case, {\tt x+7} is within the original array's bounds. In the second
case, {\tt y+7} is not within the array's bounds and the monitor will raise an
error.
